\documentclass[12pt]{article}
\usepackage[T1]{fontenc}
%\usepackage[latin9]{inputenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{setspace}
\usepackage{rotating}
\usepackage{graphics}
\usepackage[round]{natbib}
\usepackage{graphicx}
\usepackage{float} 				%allows you to float images
\usepackage{latexsym}
\usepackage{bbding}
\usepackage {moresize}
\usepackage{bbding}
\usepackage{blindtext}
\usepackage{hhline}
\usepackage{tikz}
\usetikzlibrary{shapes,backgrounds}
\usepackage{pgfplots}
\usetikzlibrary{arrows}
\usepackage{enumitem}
\doublespacing
\usepackage{geometry}
\usepackage{amsthm}
\usepackage{color}
\usepackage{array,multirow}
\usepackage{subcaption}
\usepackage{pst-plot}
	\psset{xunit=15mm}
\geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
\setlength{\parskip}{\bigskipamount}
\setlength{\parindent}{0pt}

\newenvironment{problem}[2][Problem]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}

\title{Problem Set 4 \thanks{Problem list -3.10.6, 3.10.16, 3.12.6, 3.12.8}}
\author{Ian McGroarty \\
	Course Number: 625.603}
\date{February 28, 2019}

\begin{document}

\maketitle

\begin{problem}{3.10.6} Let $Y_1, Y_2, ... , Y_n$ be a random sample from the exponential pdf $f_y(y) = e^{-y}, y \geq 0$. What is the smallest n for which $P(Y_{min} < 0.2) > 0.9$? \\
\textbf{Solution} The evaluated n > 11.513, since x must be an integer (I'm assuming since these are trials) $n\geq 12$.
\begin{align*}
P(Y_{min}<0.2) &= \int_0^{0.2} f_{Y_{min}}(y) \\
&=\int_0^{0.2} n[1-F_Y(y)]^{n-1}f_Y(y) && \text{Theorem 3.10.1.b (pg193)} \\
&=\int_0^{0.2} n[1-(1-e^{-y})]^{n-1}(e^{-y}) \\
&= \int_0^{0.2}n(e^{-y})^{n-1}(e^{-y}) \\
&= \int_0^{0.2}n(e^{-ny}) \\
&= -e^{-ny} \Big|_0^{0.2} \\
&= 1-e^{-(0.2)n} >0.9 \\
&= log(e^{-0.2n}) < log(0.1) \\
n&>\frac{log(0.1)}{-0.2} \approx 11.513
\end{align*}

\end{problem}

\begin{problem}{3.10.16} Suppose a device has thre independent components, all of whose lifetimes (in months are modeled by the exponential pdf, $f_y(y) = e^{-y}, y > 0 $. What is the probability that all three components will fail within two months of one another? 

\textbf{Solution}\footnote{To start, I want to note that understanding of the "memoryless property of the exponential distribution" was critical to even approaching success in this problem. I studied the proof in this pdf, http://www.cs.cmu.edu/afs/cs/academic/class/15750-s19/OldScribeNotes/lecture11.pdf (pg 2). I also used wolframalpha to do some of the calculations that were a to complex for my patience level.} Range = $Y_{max}-Y_{min}=Y^`_3-Y^`_1.$ The \textit{memoryless property of the exponential distribution}: $P(X\geq s+t | x \geq s) = P(X\geq t)$. This implies that the level of $Y^`_1$ is inconsequential. Thus we can assume that $Y^`_1 = 0$. In which case we are really only interested in $P(Y^`_{max} < r)$. Thus we can apply theorem 3.10.1.a (pg 193) with: n=3, $f_y(y)=e^{-y}$, and $F_Y(y) = \int_0^y f_Y(y)dy = 1-e^{-y}$. 
\begin{align*}
P(Y_{max} < m) &= \int_{-\infty}^m n[F_Y(y)]^{n-1}f_Y(y) \\
P(Y^`_3 < 2) &= \int_0^2  3[1-e^{-y}]^2 e^{-y} &&\text{Enter WolframAlpha} \\
&\approx 0.646. 
\end{align*}
\end{problem}

\begin{problem}{3.12.6} Find $M_Y(t)$ if Y has the pdf: 
$$
f_Y(y) = \left\{
        \begin{array}{ll}
            y, & \quad 0 \leq y \leq 1 \\
            2-y, & \quad 1 \leq y \leq 2 \\
		0,  & \quad \text{elsewhere}
        \end{array}
    \right.
$$

\textbf{Solution} Since X is continuous, the second part of Definition 3.12.1 (pg 206) applies, so: 
\begin{align*}
M_Y(t) &= E(e^{tW}= \int_{-\infty}^{\infty} e^{tw}f_W(w)dw  \\
&= \int_{-\infty}^{0} e^{ty}0 + \int_{0}^{1} e^{ty}y + \int_{1}^{2} e^{ty}(2-y) + \int_{0}^{\infty} e^{ty}0 \\
&= \int_{0}^{1} e^{ty}y + \int_{1}^{2} e^{ty}(2-y)  && \text{I used WolframAlpha here.} \\
&=\frac{e^{tx}(tx-1)}{t^2} \Big|_0^1 + \frac{(y-2)e^{tx}}{t}\Big|_1^2  \\
&= \frac{1-e^t}{t^2}
\end{align*}
\end{problem}


\begin{problem}{3.12.8} Let Y be a continuous random variable with $f_Y(y) = ye^{-y}, o \leq y$. Show that $M_Y(t) = \frac{1}{(1-t)^2}$ 

\textbf{Solution} Since X is continuous, the second part of Definition 3.12.1 (pg 206) applies, so: 
\begin{align*}
M_Y(t) &= E(e^{tW}= \int_{-\infty}^{\infty} e^{tw}f_W(w)dw  \\
&= \int_{0}^{\infty} e^{ty}ye^{-y}  \\
&= \int_{0}^{\infty} e^{ty-y}y  && \text{The integration by parts at the end if you want to see it.}\\
&= \frac{e^{(t-1)y}((t-1)y-1)}{(t-1)^2} \\
\end{align*}

Integration by parts
\begin{align*}
f(y) = y & \ \ \ df = dy  \\
dg(y) = e^{y(t-1)} & \ \ \ g = \frac{e^{y(t-1)}}{t-1} \\
&= \frac{y(e^{y(t-1)}}{t-1} - \frac{1}{(t-1)}\int e^{y(t-1)} \\
&= \frac{y(e^{y(t-1)}}{t-1} - \frac{e^{y(t-1)}}{(t-1)^2} \Big|_0^{\infty} \\
 \end{align*}
\end{problem}

\end{document}



 This is a matter of using joint probability to determine the pdf of the range: I use equation 3.10.5 (pg 197). With n=3, i=3, j=3, $f_Y(y) = e^{-y}$ and $F_Y(y)=-e^{-y}$.
\begin{align*}
f_{Y_i^`,Y_j^`}(u,v) &= \frac{n!}{(i-1)!(j-i-1)!(n-j)!}[F_Y(u)]^{i-1}[F_Y(v)-F_Y(u)]^{j-i-1} \cdot [1-F_Y(v)]^{n-j}f_Y(u)f_Y(v) \\
&= 6[-e^{-v}+e^{-u}](e^{-v})(e^{-u})
\end{align*}
Find $M_Y(t)$