---
title: "Problem Set 9"
author: "Ian McGroarty"
date: "03APRIL2020"
output: html_document
---


```{r setup, include=FALSE}
Sys.setenv(HTTP_PROXY = "c1proxy.frb.org:8080")
Sys.setenv(HTTPS_PROXY = "c1proxy.frb.org:8080")
Sys.setenv(SPARK_HOME = "/usr/hdp/current/spark2-client")
Sys.setenv(SPARK_CONF_DIR= "/usr/hdp/current/spark2-client/conf/")

library(knitr)
library(rmarkdown)
library(expm)
library(rmutil)
library(mnormt)
library(graphics)
library(raster)
library(sp)
library(MASS)
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1: 
Consider an atrifical dataset consisting of eight numebrs. let $\hat{\theta}$ be the 25\% trimmed mean, computer by deleting the smallest two numbers and the largets two numbers, and then taking the average of the four remaining numbers. ## (a) Calculate $\hat{Var_B}(\hat{\theta})$ for B = 25,100,00,500,1000,2000. From these results estimate the ideal bootstrap estimate $\hat{Var_{\infty}}(\hat{\theta})$

So we need to draw B samples of size n=8 from X. Order the sample from smallest to largest. Take the average of the middle four numbers. 

```{r}
X1 <- c(1,2,3.5,4,7.3,8.6,12.4,13.8)

theta <- c()
bootstrap.var <- function(DATA,B){
    for (i in 1:B){
        # Make sure the data is in vector form
          vctrzdData <- as.vector(DATA)
        # Sample of size n with replacement from data 
          btstrpSmpl = sample(vctrzdData,length(vctrzdData), replace = TRUE)
        # Sort
          btstrpSmpl.ordered <- sort(btstrpSmpl)
        # Take the mean of the middle 4 numbers 
          theta[i] <- mean(btstrpSmpl.ordered[3:6])
      }
  # Take the mean of the B thetas
    theta.mean <- mean(theta)
  # Take the variance of the B thetas (var(theta) also works)
    theta.var <- (1/(B-1))*(sum((theta-theta.mean)^2))
  return(theta.var)
}

## The cool way to do it with a line
  #X <- c()
  #for (j in 1:200){ X[j] <- bootstrap.var(DATA=X1,j)}
  #plot(X,type = 'l', xlab="Bootstrap iterations",ylab= "Variance of theta")

# Run over 10 difference seeds 
X1 <- c(1,2,3.5,4,7.3,8.6,12.4,13.8)
    for (j in c(25,100,200,500,1000,2000)){
      print(paste0("B=",j," Var(theta) = ",bootstrap.var(DATA=X1,j)))
    }
X2 <- runif(n=8,min=1,max=20)
    for (j in c(25,100,200,500,1000,2000)){
      print(paste0("B=",j," Var(theta) = ",bootstrap.var(DATA=X2,j)))
    }
X3 <- runif(n=8,min=1,max=20)
    for (j in c(25,100,200,500,1000,2000)){
      print(paste0("B=",j," Var(theta) = ",bootstrap.var(DATA=X3,j)))
    }
X4 <- runif(n=8,min=1,max=20)
    for (j in c(25,100,200,500,1000,2000)){
      print(paste0("B=",j," Var(theta) = ",bootstrap.var(DATA=X4,j)))
    }
X5 <- runif(n=8,min=1,max=20)
    for (j in c(25,100,200,500,1000,2000)){
      print(paste0("B=",j," Var(theta) = ",bootstrap.var(DATA=X5,j)))
    }
X6 <- runif(n=8,min=1,max=20)
    for (j in c(25,100,200,500,1000,2000)){
      print(paste0("B=",j," Var(theta) = ",bootstrap.var(DATA=X6,j)))
    }
X7 <- runif(n=8,min=1,max=20)
    for (j in c(25,100,200,500,1000,2000)){
      print(paste0("B=",j," Var(theta) = ",bootstrap.var(DATA=X7,j)))
    }
X8 <- runif(n=8,min=1,max=20)
    for (j in c(25,100,200,500,1000,2000)){
      print(paste0("B=",j," Var(theta) = ",bootstrap.var(DATA=X8,j)))
    }
X9 <- runif(n=8,min=1,max=20)
    for (j in c(25,100,200,500,1000,2000)){
      print(paste0("B=",j," Var(theta) = ",bootstrap.var(DATA=X9,j)))
    }
X10 <- runif(n=8,min=1,max=20)
    for (j in c(25,100,200,500,1000,2000)){
      print(paste0("B=",j," Var(theta) = ",bootstrap.var(DATA=X10,j)))
    }

```


# Problem 2
  Patients with advanced terminal cancer of the breast were treated with ascorbate in an attempt to prolong survival. 

## (a) Construct a 95\% confidence interval 
of the the mean breast cancer survival time by apply the simple bootstrap to logged data and exponentiating the resutling interval boundaries. 
  
```{r}
B <- 200 

Survival = c(25, 42, 45, 46, 51, 103, 124, 146, 340, 396, 412, 879, 1112)
survival.log <- log(Survival)

## DONT FORGET TO EMPTY THETA 
  theta <- c()

    for (i in 1:200){
        # Make sure the data is in vector form
          vctrzdData <- as.vector(survival.log)
        # Sample of size n with replacement from data 
          btstrpSmpl = sample(vctrzdData,length(vctrzdData), replace = TRUE)
        # Take the mean
          theta[i] <- mean(btstrpSmpl)
      }
  # Take the mean of the B thetas
    theta.mean <- mean(theta)
  # 95% confidence interval for 200 observations mean the 5th and 195th observations
    theta.ordered <- sort(theta)
    print(paste0(
      "The 95% confidence interval for the logged survival data with 200 bootstrap iterations is " 
      ,exp(theta.ordered[c(5)]),",",exp(theta.ordered[c(195)])))
```
  
## (b) Construct another 95% using original data
```{r}
B <- 200 

Survival = c(25, 42, 45, 46, 51, 103, 124, 146, 340, 396, 412, 879, 1112)
#survival.log <- log(Survival)

## DONT FORGET TO EMPTY THETA 
  theta <- c()
    for (i in 1:B){
        # Make sure the data is in vector form
          vctrzdData <- as.vector(Survival)
        # Sample of size n with replacement from data 
          btstrpSmpl = sample(vctrzdData,length(vctrzdData), replace = TRUE)
        # Take the mean
          theta[i] <- mean(btstrpSmpl)
      }
  # Take the mean of the B thetas
    theta.mean <- mean(theta)
  # 95% confidence interval for 200 observations mean the 5th and 195th observations
    theta.ordered <- sort(theta)
    print(paste0(
      "The 95% confidence interval for the raw survival data with 200 bootstrap iterations is " 
      ,theta.ordered[c(5)],",",theta.ordered[c(195)]))
```
  
# Problem 3
Assume $X_1 \cdots X_n \ \sim Unif(0,\theta)$. What is the maximum likelihood estimate of $\theta $. 

For reference: (Marx \& Larsen (2018), p. 281). This is a strange question since the uniform distribution will have an equal probablity $(p_X = \frac{1}{\theta})$,  for all numbers in the support of $X_n$. That means that the maximum likelihood estimate of $\theta = \theta$. Or more formally, 
$$\begin{align*}
L(\theta) &= \prod_{i=1}^{n} \frac{1}{\theta} = \theta^{-n} \\
\frac{d ln L(\theta)}{d \theta} &= \frac{d}{d\theta}-nln(\theta) = \frac{-n}{\theta} \\
\theta &= \theta_e =x_{max}
\end{align*}$$

So the question becomes, how do estimate $\theta$ if it is simply the max of the sample? 


# Problem 4
Suppose $X^*= (x_1^*,\cdots,x_n^*)$ is a bootstrap sample obtained from $x=(x_1,\cdots,x_n)$. Show that the probability that any particular value of x is in $x^*$ exactly k times. 

So we have a set $X=(x_1,\cdots,x_n)$ and we draw a sample with replacement from x to obtain  $X^*= (x_1^*,\cdots,x_n^*)$. This is equvalent to drawing a uniform random variable from the support of X. Thus, each element of the set x is drawn with probability $\frac{1}{n}$. 

Given $x_j \in X$ and $x^*_i \in X^*$, let $x_1^* \cdots x_n^* $ be n independent trials, each resulting in either success or failure, with success defined as $x_i^* = x_j$. The probablity of k success is a binomial random variable with the probability of success for any trial $p = \frac{1}{n}$. Therefore, by Theorem 3.2.1 (Larsen & Marx (2018), p. 104):
$$\text{P(k successes)}={n \choose k}(\frac{1}{n})^k (\frac{n-1}{n})^{n-k} $$

We can see this in R by calculating the theoretical probability and creating a bootstrap sample for $$\hat{\theta}_i^* = \sum_{i=1}^nI_{(x_j^* = x_i)} $$
Where $I(\phi)$ is a Binomial random variable with success probability $\phi $

```{r}
## Define nchoosek (Larsen & Marx (2018), p. 84). 
  nchoosek <- function(n,k){
    factorial(n)/((factorial(k))*(factorial(n-k)))
  }

## Define full function (Larsen & Marx (2018), p. 104)
 binom.nk <- function(n,k){
   nchoosek(n,k) * ((1/n)^k) * ((n-1)/n)^(n-k)
 }
 

## Get a base X of size n
 vctrzdData <- c(1,2,3.5,4,7.3,8.6,12.4,13.8)

## Set parameters
 n <- length(vctrzdData)
 k <- 2
 B <- 1000

## Boot Strap 
  count <- c()
    for (i in 1:B){
      ## Get B samples from X
        btstrpSmpl = sample(vctrzdData,length(vctrzdData), replace = TRUE)
      ## Count how many times x_i shows up in the sample
        count[i] <- sum(btstrpSmpl == vctrzdData[3])
    }
## Probability
  (sum(count == k))/B

## Compare to theoretical probability 
  binom.nk(length(vctrzdData),k)
```

# Problem 5
We know that $\hat{\theta}=\bar{X}$ is an unbiased estimate of the mean for a particular density. So the true bais, $Bias_{True}(\hat{\theta})=0.$. Prove that 

Just for the record, the notation here is really throwing me off. 

## (a) $Bias_{True}(\hat{\theta})=0$

If $\hat{\theta}=\bar{X}$ is an unbiased estimate for $E(\theta)$ then by Definition 5.4.1 (Larsen \& Marx (2018) p. 310) that $E(\hat{\theta}) = \theta$. Thus, 
$$bias(\hat{\theta}) = E(\hat{\theta})- \theta = \theta - \theta = 0 $$

## (b) $\hat{Bias}_{Revised}(\hat{theta})= 0$
$$\begin{align*}
\hat{Bias}_{revised}(\hat{\theta}) &=
\hat{\theta}^*(\cdot)-\hat{\theta}_{revised} \\
&= \frac{1}{B}\sum_{j=1}^B\hat{\theta^*_i} - g(\bar{P}^*x^T) \\ 
&= \frac{1}{B}\sum_{j=1}^Bg(x^*_j) - g(\frac{1}{B}\sum_{j=1}^BP^*_jx^T) \\ 
\text{As } lim_{B\rightarrow \infty}P^*_j \rightarrow (\frac{1}{n}) &\implies P^*_jx^T = x_j\cdot(\frac{1}{n}) \\
\end{align*}$$

Since $x_j$ is continuously sampled from the empirical distribution function, these two terms are equvalent. 


## (c) show that it's not necessarily the case that $\hat{Bias}(\hat{\theta})=0.$

Since we have shown that $\hat{\theta}^*(\cdot)=\hat{\theta}_{revised}$ it suffices to show that $\hat{\theta}_{revised} \neq \hat{\theta}$ Which was shown in problems one and two which showed that $\hat{\theta}$ had non-zero variance and depended heavily on the size of B. So while they may be assymtotically equal, they are not equal by construction. 


```{r, include=FALSE}


bootstrapExample = function(B)
{
	#vctrzdData = as.vector(presidents);
	#vctrzdData = vctrzdData[!is.na(vctrzdData)];
	btstrpMeans = rep(0,B);
	for (j in 1:B) {
		btstrpSmpl = sample(vctrzdData,length(vctrzdData), replace = TRUE);
		btstrpMeans[j] = mean(btstrpSmpl);
	}
	hist(btstrpMeans, breaks = 20);
	print("Bootstrap Standard Error");
	btstrpStdErr = sd(btstrpMeans);
	print(btstrpStdErr);
	print("Regular Standard Error");
	regStdErr = sd(vctrzdData)/sqrt(length(vctrzdData));
	print(regStdErr);
	quantile(btstrpMeans, c(.025, .975));
	
}	

bootstrapExample(10000)
```

