---
title: "MonteCarlo"
author: "Ian McGroarty"
date: "February 24, 2020"
output: html_document
---

```{r setup, include=FALSE}
#library(knitr)
#library(rmarkdown)
library(expm)
#library(matlib3)
#knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 2

## Distributions 
```{r}
# Three independent generaTIONS FROM A GAMMA DISTRIBUTION WITH mean = 2.5 and variance = 4.5
  rgamma(3,2.5,4.5)
# Uniform - 100 independent generations from a uniform distribution with min 2 and max 5
  runif(100, min=2, max = 5)
# Gamma distributions?
  # g1 <- rgamma(10000,6,rate=(p/(1-p)))
  # hist(g1, freq=FALSE)
  # g2 <- rgamma(n=10000,shape=6,rate=(p/(1-p)))
  # hist(g2, freq=FALSE)
  # g3 <- rgamma(n=10000,shape = 6, scale = ((1-p)/p))
  # hist(g3, freq=FALSE)
  
  
Nsim <- 5000
n <- 4.3
p <- 6.2
### NOTE THE BOOK PARAMETERIZES GAMMA AS G ~ (n,B) where n is the mean (lambda) and B = ((1-p)/p). Where B is the rate parameter and p is probabilities. I'm a little confused here so I'll need to come back to this. Is it scale or rate?
y <- rgamma(Nsim,shape=n,scale=6.2)
hist(y,freq=FALSE)  


  
```


```{r cars}
Nsim <- 10^4      # number of random variables
x <- runif(Nsim)
x1 <- x[-Nsim]    # Vectors to plot?
x2 <- x[-1]
par(mfrow=c(1,3)) #??
hist(x)
plot(x1,x2)
#acf(c)
```

## 2.2 Inverse Transform Method
If $X \sim Exp(1)$ then $F(X) = 1 - e^{-x}$

```{r pressure, echo=FALSE}
Nsim <- 10^4
U <- runif(Nsim)
X <- - log(U)     # Log of the uniform random variables
Y <- rexp(Nsim)   # Exponentials from the real funciton
par(mfrow=c(1,2))
hist(X,freq=F, main = "Exp from uniform")
hist(Y,freq=F, main = "Exp from Funciton")
```


### 2.2.2 General Transformation methods

To generate $\chi^2$ random variables:
```{r}
U <- runif(3*10^4)
## nrow *2 = df
U <- matrix(data=U, nrow=6)
X <- -log(U)
X <- 2 * apply(X,2,sum)
y <- rchisq(U,df=12)
hist(X,freq=F, main = "Exp from Uniform")
hist(y,freq=F, main = "Exp from ChiSq")
```

### 2.2.2 Discrete Distributions
#### Example 2.5 
Generate Poisson random variable for large values of lambda. 
```{r}
Nsim <- 10^4
lambda <- 100
spread <- 3*sqrt(lambda)
t <- round(seq(max(0,lambda-spread),lambda+spread,1))
prob <- ppois(t,lambda) # get the poisson probability for all values in the range 
X <- rep(0,Nsim)# Create a space for the values to go
for (i in 1:Nsim){
  u <- runif(1)
  X[i] <- t[1] + sum(prob<u) } 
# Sum all of the values of prob that are less than U 
# This assigns the poisson probability to all of the numbers 
hist(X, freq=F)
Y <- rpois(1000,100)
hist(Y, freq=F)
```
## 2.3 Accept Reject Methods
The Upper bound M is then the maximum of the beta density 
```{r}
## Obtain M the maximum of the distibution
M <-  optimise(f=function(x){dbeta(x,2.7,6.3)},interval=c(0,1),maximum=T)$objective
a <- 2.7
b <- 6.3
U <- runif(Nsim, max = M)
y <- runif(Nsim)
x <- y[U < dbeta(y,a,b)]
hist(x)

```

### Accept reject with Normal from exam 1
```{r}

#### Exam Problem 8 ####
Nsim <- 10000

## Set Up  
  # Set the function  
    fx <- function(x){ 
      30*(x^2-2*x^3+x^4) 
    }
  # Find the Max
    Xcand <- runif(Nsim,min=0,max=1)
    M <- max(fx(Xcand))
    
  # Get a look to see what it is like 
    plot(fx,xlim=c(0,1),ylim=c(0,2.5),col="blue")
    
## Approx with Uniform Distribution
    
    ## Approx with Uniform Distribution
    # X candidates from Uniform
    Xcand <- runif(Nsim,min=0,max=1)
    # Generate y candidates up to M
    Ycand <- runif(Nsim, min=0, max=M)
    # Keep the x candidates for which the Y candidates are viable solutions
    XB <- Xcand[Ycand < fx(Xcand)]
    # Acceptence Rate
    length(XB)/Nsim
    # Plot to see how close we are
    hist(XB,freq=F,ylim=c(0,2),main="Hist via Acc/Rej with Uniform",add=FALSE)
    plot(fx,xlim=c(0,1),ylim=c(0,2),col = "blue",add=TRUE)
    
    
    ## Get a scatter plot
    DF <- as.data.frame(Xcand)
    DF$Ycand <- runif(Nsim, min=0, max=M)
    DF$Acc <-ifelse(DF$Ycand < fx(DF$Xcand),0,1)
    
    ggplot(DF,aes(x=DF$Xcand,y=DF$Ycand,color=DF$Acc)) + geom_point(size=2) + 
      ggtitle("Scatter Acc/Rej with Uniform") +
      stat_function(fun=function(x)(30*(x^2-2*x^3+x^4)),geom="line",col="red",size=2 )
    
## Approx With Normal Distribution
    # Find the Approriate mean and variance for the cumulative normal distribution
      gx <- function(x){ 
        (dnorm(x,0.5,0.2)) 
      }
      
      plot(fx,xlim=c(0,1),ylim=c(0,2.5),col="blue")
      plot(gx,col="red",add = TRUE)
    
    # Find normalizing constant s.t. g(x)*M > f(x) for all x 
      ## Set hx = fx/gx
        hx <- function(x){
          fx(x)/gx(x)
        }
      ## Find the Max of hx(x) 
        hx.opt <- optimise(hx,interval =c(0,1),maximum=TRUE)$maximum
      ## Get the Constant M = hx(hx.opt) => fx < gx*M  
        M <- hx(hx.opt)   
    
      # Redefine gx do include the M 
        gx2 <- function(x){ 
          M*(dnorm(x,0.5,0.2)) 
        }
        
      # See that indeed gx*M > fx for all X 
        plot(fx,xlim=c(0,1),ylim=c(0,2.5),col="blue")
        plot(gx2,col="red",add = TRUE)
        
    # X Candidates from Normal 
      ## I can't figure out how to set a range with the rnorm command so this is what I came up with 
        counter <- 0
        Xcand <- c()
        while(counter<Nsim) {
          z <- 1*rnorm(1,0.5,0.2)
          if (z <= 1 & z>=0){
            Xcand <- append(Xcand,z)
            counter = counter + 1 
          } 
        }
    # Y Candidates from Uniform up to M*gx(Xcand)
      Ycand <- runif(Nsim,min=0,max=(gx2(Xcand)))
    # Accept or Reject
      XB <- Xcand[(Ycand < (fx(Xcand)))]
    # Accept Rate  
      length(XB)/Nsim
    # See how Close it is 
      hist(XB,freq=F,ylim=c(0,2.5),breaks = 30,main="Hist via Accept Reject")
      plot(fx,xlim=c(0,1),ylim=c(0,2),col = "blue",add=TRUE)
      
      
    ## Get a scatter Plot
      DF <- as.data.frame(Xcand)
      DF$Ycand <- runif(Nsim,min=0,max=gx2(Xcand))
      DF$Acc <-ifelse(DF$Ycand < fx(Xcand),1,0)
      sum(DF$Acc)/Nsim
      
      
      ggplot(DF,aes(x=DF$Xcand,y=DF$Ycand,color=DF$Acc)) + geom_point(size=2) + ylim(0,2) +
        stat_function(fun=function(x)(30*(x^2-2*x^3+x^4)),geom="line",col="red",size=2 )
      
```


# Chapter 3 Monte Carlo Integration

### Exercise 3.1
```{r}
f1 <- function(t){ 
  t/(1+t*t)*exp(-(0-t)^2/2)
  }
f2 <- function(t){ 
  1/(1+t*t)*exp(-(0-t)^2/2)
  }
curve(f1,-3,3,col=1,ylim=c(-0.5,1),xlab="t",ylab="",ty="l") 
curve(f2,-3,3,col=2,ty="l",add=TRUE) 

```



### Example 3.3 
Evaluate the integral from [0,1] of $h(x)=[cos(5ox)+sin(20x)]^2$

```{r}
## Set the function
  h <- function(x){
    (cos(50*x)+sin(20*x))^2
  }

## Integrate
  truemean <- integrate(h,0,1)

## Sampling
  # Set sim 
    Nsim <- 10^4
  # Generate uniform
    x <- runif(Nsim)
  # Evaluate h(x)
   hx <- h(x) 
  # Average of hx
   estint <- cumsum(hx)/(1:Nsim)
  # Get the variation from the true mean for everyvalue of hx
   esterr <- sqrt(cumsum(hx-estint)^2)/(1:Nsim)

## Graph
  #par(mar=c(2,2,2,1),mfrow=c(1,2))
  #curve(h,xlab="Function",ylab="",lwd=2)
  plot(estint, xlab="Mean and Error range",ylab="",type='l',lwd=2,
       ylim=mean(hx)+20*c(-esterr[Nsim],esterr[Nsim]))
  
  lines(estint+2*esterr,col="gold",lwd=1)
  lines(estint-2*esterr,col="gold",lwd=1)
  

```

### Example 3.4
We can caluclate the cumulative normal table using monte carlo simluation:
```{r}
# Generate values from normal distribution:
  x <- rnorm(10^8)

# Set the probability values you want to calculate
  bound <- qnorm(c(.5,.75,.8,.9,.95,.99,.9999))

# 
  res <- matrix(x,ncol=8,nrow=7)
# loop for each bound and for each sample size:
  for (i in 2:8) {
  for (j in 1:8) {
    res[i-1,j]=mean(x[1:10^i]<bound[j]) 
  }
  }
  #matrix(as.numeric(format(res,digi=4)),ncol=8)

```

# Chapter 6
## ## Section 6.2 A peak at Markov Chain Theory
### Example 6.1 
If the candidate density is Uniform it doesn't depend on the pervious value of the chain so we don't need to include the q probability. 
```{r}
# Initial Values
  a <- 2.7
  b <- 6.3
  c <- 2.669 
  Nsim <- 5000
  y1 <- 1:Nsim   #for plots
  
# Initialie the chain (X CANDIDATES FROM A UNIFORM DISTRIBUTION)
   X <- rep(runif(1),Nsim)  
   
# Run the Chain
   for (i in 2:Nsim){
     Y <- runif(1)  # Canidate values of X[i+1]
     rho <- dbeta(Y,a,b)/dbeta(X[i-1],a,b)
     ## If runif(1)<rho then X[i]=X[i-1] if not it is Y
     X[i] <- X[i-1] + (Y-X[i-1])*(runif(1)<rho)
   }
   

## Plot the iterations 
  plot(y1,X,type="S", xlim=c(100,300))   
  
## Plot Beta
  B <- rbeta(Nsim,2.7,6.3)
  par(mfrow = c(1,2))
  hist(B, freq=FALSE)
  hist(X)

```


# Chapter 7 Gibbs
### Example 7.2

```{r}

Nsim = 5000
n = 15
a = 3
b= 7
## Set up the inital Array
  X <- T <- array(0,dim=c(Nsim,1))  # Initial Array
  T[1] <- rbeta(1,a,b)              # Initial Chain?
  X[1] <- rbinom(1 , n,T[1])


## Do the Distribution
  for (i in 2:Nsim){
    X[i] <- rbinom(1,n,T[i-1])
    T[i] <- rbeta(1,a+X[i],n-X[i]+b)
    }

# True distribution of Theta (T)
fx <- function(x){
  dbeta(x,a,b)
}

gx <- function(x){
  dbinom(x,n,T)
}

hist(X,freq=F)
#plot(gx,col="blue",add=TRUE)

hist(T,freq=F)
plot(fx,col="blue",add=TRUE)

```

### Example 7.5
a. the (normal) full conditionals are defined in Example 7.4. An R program that implements this Gibbs sampler is
$$ 
X_i | x_{-i)} \sim N(\frac{(p-1)\rho}{1 + (p-2)\rho)}x_{-i},\frac{1+(p-2)\rho - (p-1)\rho^2}{1+(p-2)\rho})
$$
```{r}
# Set up the parameters as in the question
  T <- 500
  p <- 5
  r <- 0.25

# Get the starting distributions 
  ## X should have p samples from the normal distribution
    X <- rnorm(p)
    
  ## Cur should be x 
    itteration <- X
# Run the sampling 
  for (t in 1:T){
    for (j in 2:p) {
      avgx <- sum(itteration[j-1])/(p-1)
      itteration[j] <- rnorm(1,(p-1)*r*avgx/(1+(p-2)*r),sqrt((1+(p-2)*r-(p-1)*r^2)/(1+(p-2)*r)))
    }
    X <- rbind(X,itteration)
  }


        
# Plots
  for (i in 1:p){
    hist(X[,i],freq=FALSE)
    curve(dnorm(x),add=TRUE)
  }

```

